Discrepancy, Coresets, and Sketches in Machine Learning

### 2019-05-09 response.md

### Program Committee Member1

The review is very thoughtful and thorough, thank you. We would only like to point out that the presented result actually does give generalization bounds. In fact, class discrepancy of sqrt(d)/m immediately translates to generalization of the form 
R < R* + O(sqrt(d)/m)
Where R is the Risk of the model obtained by ERM on an optimal coreset of size m.
The authors thought it was not worth including in paper but would be happy to include in the final version.

### Program Committee Member3

* We indeed depend on a scaling parameter unlike the result of Philips et al. Still, the contribution is a new view and the application of kernel density estimation is one of many
* Corollaries 17,18 follow from a trivial bound to the coefficient in the Taylor expansion of the functions they apply to. We will note this in the paper
* Theorem 16 indeed "plays with notations" mainly in order to make the vector balancing lemma (14) standalone. It applies for compact sets, in particular of radius R. We will make this clearer
* The improved merge-and-reduce hold for *all* queries exactly like previous constructions. The analysis goes through a single query high probability bound and union bound which might be the source of confusion.

@@

*Does prior work assume gamma to be constant?*

Some do not. This is a side effect of using a Taylor series based argument which is arguably too weak. We hope to eliminate that limitation in future research.

*How does the paper compare to earlier work? What are the novel ideas?*

The core of the paper is based on three contribution all of which are new
1. A new measure of complexity for function classes, the class discrepancy.
2. New generalization results, coresets, and streaming algorithms based only on class discrepancy. Offline coresets constructions are indeed trivial and provided for completeness (which is clearly stated and not "over sold"). The streaming algorithms are new and improve on prior results.
3. Tight bound on the class discrepancy of several well studied problems. This requires some mathematically involved lemmas independently useful in generalizing our results.

### 2019-05-09 komlos.tex

Naturally, one could iterate this process. Starting with $n$ items and ending with $k$.
Let $F_t$ denote the (unweighted) sum of functions $f$ after $t$ iterations and $n_t$ denote the cardinality of the coreset\footnote{We use the term $a=b \pm c$ to mention $|a-b| \leq c$ }. 
\[
F = F_0 = 2F_1 \pm n_0D_{n_0} = 4F_{2} \pm 2n_1D_{n_1} \pm n_0D_{n_0} = \ldots = 2^T F_T \pm \sum_{t=0}^{T-1} 2^t n_t D_{n_t}
\]
Here $T$ stand for the total number of iterations. Let us analyze the error term.
Given $n_t \le n/2^{t}$ and the polynomial dependence of $D_m$ on $m$ we have 
$$\sum_{t=0}^{T-1} 2^t n_t D_{n_t} \le n \sum_{t=0}^{T-1} D_{n/2^t} = n\cdot O(D_{n_T}) . $$ 
Setting $n_T$ in a way that $D_{n_T} = O(\eps)$ leads to a coreset with item weights of $w^T$ of the appropriate cardinality and completes the proof.

@@

\subsection{Towards an Efficient Algorithm}

%\section{A Simple Algorithm for Kernel Density Estimation}
From section \ref{sec:analytic} we know that the class discrepancy of the Gaussian kernel is $D_m = O(\sqrt{d}/m)$. 
Here, we provide a computationally efficient bound that can be achieved with a straightforward algorithm of complexity $O(m^2)$. Together with the results of Section~\ref{sec:sketch} this provides an efficient sketching algorithm for Kernel Density Estimation. 
In fact, we show that for any positive kernel $D_m = O(1/\sqrt{m})$. This bound is superior to that of the previous section for high dimensions $d > m$. 
More importantly, there is a very simple, intuitive, and deterministic algorithm for computing the signs $\sigma$. 
Given a collection of data points $X = \{x_1,\ldots, x_n\}$ in $\R^d$ the density function $f: \R^d \rightarrow \R$ of a point $q$ is defined as $ F(q) = \sum_{i=1}^{n} K(x_i,q) $.
Here, $K$ is any \emph{positive semi-definite kernel} function. The most frequent examples include
$$ K(x,q) = \exp(- \|x-q\|_2^2/\lambda^2),\;\; K(x,q) = \exp(- \|x-q\|/\lambda) ,\;\;  K(x,q) = (1+\|x-q\|/_2^2/\lambda^2)^{-1}$$
where $\lambda$ is a scaling parameter. For simplicity, we assume that $K(x,x) \leq 1$ for all data points. Notice that for any kernel based on the distance we have $K(x,x)=1$ exactly for all $x \in \R^d$.

>>>>

Naturally, one could iterate this process. Starting with $n$ items and ending with $m$.
Let $F_t$ denote the (unweighted) sum of functions $f$ after $t$ iterations and $n_t$ denote the cardinality of the coreset\footnote{The expression $a=b \pm c$ means $|a-b| \leq c$}. 
\[
F = F_0 = 2F_1 \pm n_0D_{n_0} = 4F_{2} \pm 2n_1D_{n_1} \pm n_0D_{n_0} = \ldots = 2^T F_T \pm \sum_{t=0}^{T-1} 2^t n_t D_{n_t}
\]
Here $T$ stand for the total number of iterations. Let us analyze the error term.
Given $n_t \le n/2^{t} \approx m$ and the polynomial dependence of $D_m$ on $m$ we have 
$$\sum_{t=0}^{T-1} 2^t n_t D_{n_t} \le n \sum_{t=0}^{T-1} D_{n/2^t} = n\cdot O(D_{m}) . $$ 
Setting $m$ for which $D_{m} = \eps$ gets coresets with appropriate cardinalities and completes the proof.

@@

\subsection{Towards an Efficient Algorithm}\label{kde2}
%\section{A Simple Algorithm for Kernel Density Estimation}

From section \ref{sec:analytic} we know that the class discrepancy of the Gaussian kernel is $D_m = O(\sqrt{d}/m)$. 
Here, we provide a computationally efficient bound that can be achieved with a straightforward algorithm of complexity $O(m^2)$. Together with the results of Section~\ref{sec:sketch} this provides an efficient sketching algorithm for Kernel Density Estimation. 
In fact, we show that for any positive kernel $D_m = O(1/\sqrt{m})$. This bound is superior to that of the previous section for high dimensions $d > m$. 
More importantly, there is a very simple, intuitive, and deterministic algorithm for computing the signs $\sigma$. 
Given a collection of data points $X = \{x_1,\ldots, x_n\}$ in $\R^d$ the density function $f: \R^d \rightarrow \R$ of a point $q$ is defined as $ F(q) = \sum_{i=1}^{n} K(x_i,q) $.
Here, $K$ is any \emph{positive semi-definite kernel} function. The most frequent examples include
$$ K(x,q) = \exp(- \|x-q\|_2^2/\lambda^2),\;\; K(x,q) = \exp(- \|x-q\|/\lambda) ,\;\;  K(x,q) = (1+\|x-q\|/_2^2/\lambda^2)^{-1}$$
where $\lambda$ is a scaling parameter. For simplicity, we assume that $K(x,x) \leq 1$ for all data points. Notice that for any kernel based on the distance we have $K(x,x)=1$ exactly for all $x \in \R^d$.
